---
title: "P8105 Homework 2 - Tanya Butt (thb2114)"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

I will import the "Mr. Trash Wheel" dataset, clean the variable names, remove
non-specific dumpster data, and round the number of sports balls to the nearest integer.

```{r, collapse = TRUE}
mr_trash_wheel_df = read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                               range = "A2:N408")
mr_trash_wheel_df = janitor::clean_names(mr_trash_wheel_df) 
mr_trash_wheel_df = drop_na(mr_trash_wheel_df, dumpster) %>% 
  mutate(sports_balls = round(sports_balls, digits = 0))
```

I will now clean and combine the precipitation data for 2019 and 2018. Rows without precipitation data 
were initially omitted, a year variable was added, and the month variable was converted to a character variable.

```{r, collapse = TRUE}
precipitation_2019_df = read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 4,
                               range = "A2:B14") 
precipitation_2019_df = drop_na(precipitation_2019_df, Total) %>% 
                        add_column(Year = "2019")

precipitation_2018_df = read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 5,
                               range = "A2:B14") 
precipitation_2018_df = drop_na(precipitation_2018_df, Total) %>% 
                        add_column(Year = "2018")

precipitation_2018_2019_df = bind_rows(precipitation_2018_df, precipitation_2019_df) %>% 
  janitor::clean_names() %>% 
  pivot_wider(
    names_from = "year",
    values_from = "total") %>% 
  mutate(month = month.name[as.numeric(month)])

nrow(precipitation_2018_df)
nrow(precipitation_2019_df)
```

The overall "Mr. Trash Wheel" dataset includes the following variables: `r names(mr_trash_wheel_df)`. It has `r ncol(mr_trash_wheel_df)` columns and `r nrow(mr_trash_wheel_df)` observations.

Monthly and yearly precipitation amounts are also recorded in sheets in the "Mr. Trash Wheel" dataset. The variables included in these sheets are: `r names(precipitation_2018_df)`. 

In 2018, there were `r nrow(precipitation_2018_df)` months of precipitation. In 2019, only `r nrow(precipitation_2019_df)` months of precipitation were recorded.

```{r}
total_precip_2018 = sum(pull(precipitation_2018_2019_df, `2018`))
```

For 2018, the total amount of precipitation was `r total_precip_2018` in.

```{r}
mr_trash_wheel_df = select(mr_trash_wheel_df,year, sports_balls) %>% 
  filter(year == "2019")

median_sports_balls_2019 = median(pull(mr_trash_wheel_df, sports_balls))
median_sports_balls_2019 = round(median_sports_balls_2019, digits = 0)
```

The median number of sports balls in a dumpster in 2019 was `r median_sports_balls_2019`.
