---
title: "P8105 Homework 2 - Tanya Butt (thb2114)"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
```

## Problem 1

I will import the "Mr. Trash Wheel" dataset, clean the variable names, remove
non-specific dumpster data, and round the number of sports balls to the nearest integer.

```{r, collapse = TRUE}
mr_trash_wheel_df = read_excel("Trash-Wheel-Collection-Totals-new.xlsx", 
                               range = "A2:N535")
mr_trash_wheel_df = janitor::clean_names(mr_trash_wheel_df) 
mr_trash_wheel_df = drop_na(mr_trash_wheel_df, dumpster) %>% 
  mutate(sports_balls = round(sports_balls, digits = 0))
```

I will now clean and combine the precipitation data for 2019 and 2018. Rows 
without precipitation data were initially omitted, a year variable was added,
and the month variable was converted to a character variable.

```{r, collapse = TRUE}
precipitation_2019_df = read_excel("Trash-Wheel-Collection-Totals-new.xlsx",
                                   sheet = 6,range = "A2:B14") 
precipitation_2019_df = drop_na(precipitation_2019_df, Total) %>% 
                        add_column(Year = "2019")

precipitation_2018_df = read_excel("Trash-Wheel-Collection-Totals-new.xlsx",
                                   sheet = 7, range = "A2:B14") 
precipitation_2018_df = drop_na(precipitation_2018_df, Total) %>% 
                        add_column(Year = "2018")

precipitation_2018_2019_df = bind_rows(precipitation_2018_df, 
                                       precipitation_2019_df) %>% 
  janitor::clean_names() %>% 
  pivot_wider(
    names_from = "year",
    values_from = "total") %>% 
  mutate(month = month.name[as.numeric(month)])

nrow(precipitation_2018_df)
nrow(precipitation_2019_df)
```

The overall "Mr. Trash Wheel" dataset includes the following variables: `r names(mr_trash_wheel_df)`. It has `r ncol(mr_trash_wheel_df)` columns and `r nrow(mr_trash_wheel_df)` observations.

Monthly and yearly precipitation amounts are also recorded in sheets in the "Mr. Trash Wheel" dataset. The variables included in these sheets are: `r names(precipitation_2018_df)`. 

In 2018, there were `r nrow(precipitation_2018_df)` months of precipitation. In 2019, `r nrow(precipitation_2019_df)` months of precipitation were also recorded.

```{r}
total_precip_2018 = sum(pull(precipitation_2018_2019_df, `2018`))
```

For 2018, the total amount of precipitation was `r total_precip_2018` in.

```{r}
mr_trash_wheel_df = select(mr_trash_wheel_df,year, sports_balls) %>% 
  filter(year == "2019")

median_sports_balls_2019 = median(pull(mr_trash_wheel_df, sports_balls))
median_sports_balls_2019 = round(median_sports_balls_2019, digits = 0)
```

The median number of sports balls in a dumpster in 2019 was `r median_sports_balls_2019`.

## Problem 2

I will merge 3 datasets from FiveThirtyEight into a single dataset using year and month.

```{r, collapse = TRUE, message = FALSE}
pols_month_df = read_csv("pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, c("year", "month", "day")) %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(month = month.abb[as.numeric(month)], 
         month = str_to_lower(month)) %>% 
  mutate(day = as.numeric(day)) %>% 
  mutate(prez_gop = recode(prez_gop, `0` = "dem", 
                           `1` = "gop", `2` = "gop")) %>% 
  select(year, month, prez_gop, gov_gop, sen_gop, rep_gop, gov_dem, sen_dem, 
         rep_dem) %>% 
  rename(president = prez_gop)

snp_df = read_csv("snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(date = mdy(date)) %>% 
  separate(date, c("year", "month", "day")) %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(year = ifelse(year > 2021, year - 100, year)) %>% 
  mutate(month = as.numeric(month)) %>% 
  mutate(month = month.abb[as.numeric(month)], month = str_to_lower(month)) %>% 
  relocate(year, month, day, close) %>%
  rename(s_p_closing_value = close) %>% 
  select(year, month, s_p_closing_value)

unemployment_df = read_csv("unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
  jan:dec,
   names_to = "month",
   values_to = "unemployment_percentage")

snp_pols_df = left_join(pols_month_df, snp_df, by = c("year" = "year",
                                                      "month" = "month"))
snp_pols_unemp_df = left_join(snp_pols_df, unemployment_df, by = c("year" = "year", "month" = "month"))

```

The dataset "pols month" contains `r nrow(pols_month_df)` observations for `r ncol(pols_month_df)` variables related to the number of politicians who are democratic or republican from 1947 to 2015. The variable names in this dataset include: `r names(pols_month_df)`.

The dataset "snp" contains `r nrow(snp_df)` observations for `r ncol(snp_df)` variables related to the Standard and Poor's (S&P) stock market index. The years included range from 1950 to 2015. The variable names in this dataset include: `r names(snp_df)`.

The dataset "unemployment" contains `r nrow(unemployment_df)` observations for `r ncol(unemployment_df)` variables related to the monthly percentage of unemployment from 1950 to 2015. The variable names in this dataset include: `r names(unemployment_df)`.

The combined version of these datasets include the following variables: `r names(snp_pols_unemp_df)`

## Problem 3

I will now load the popular baby names dataset from NYC Open data. I will rename the gender predictor, recode the ethnicity variable, and remove duplicate rows.
```{r}
babynames_df = read.csv("Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  rename(sex = gender) %>% 
  mutate(ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
                            "BLACK NON HISP" = "BLACK NON HISPANIC", 
                            "WHITE NON HISP" = "WHITE NON HISPANIC")) %>% 
  mutate(child_s_first_name = toupper(child_s_first_name)) %>% 
  distinct(year_of_birth, sex, ethnicity, child_s_first_name, count, rank)
```

I will now create a table showing the rank in popularity and name count of the female baby name "Olivia" over time.
```{r}
olivia_df = filter(babynames_df, child_s_first_name == "OLIVIA") %>% 
  pivot_wider(names_from = year_of_birth, values_from = c(rank,count),
              values_fill = 0) %>% 
  relocate(child_s_first_name, sex, ethnicity, rank_2011, count_2011, rank_2012, 
           count_2012, rank_2013, count_2013, rank_2014, count_2014,
           rank_2015, count_2015, rank_2016, count_2016) %>% 
  rename("Child's Name" = child_s_first_name) %>% 
  rename(Sex = sex) %>% 
  rename("Race/Ethnicity" = ethnicity) %>% 
  rename("Rank in 2011" = rank_2011) %>% 
  rename("Count in 2011" = count_2011) %>% 
  rename("Rank in 2012" = rank_2012) %>% 
  rename("Count in 2012" = count_2012) %>%
  rename("Rank in 2013" = rank_2013) %>% 
  rename("Count in 2013" = count_2013) %>% 
  rename("Rank in 2014" = rank_2014) %>% 
  rename("Count in 2014" = count_2014) %>%   
  rename("Rank in 2015" = rank_2015) %>% 
  rename("Count in 2015" = count_2015) %>% 
  rename("Rank in 2016" = rank_2016) %>% 
  rename("Count in 2016" = count_2016) %>% 
  knitr::kable()

olivia_df
```

I will now produce a table ranking the most popular name among male children over time.

```{r, collapse = TRUE}
#Identify male names with rank = 1
popular_male_names_df = select(babynames_df, year_of_birth, sex, ethnicity,
                                child_s_first_name, rank) %>% 
  filter(sex == "MALE", rank == '1') 

popular_male_names_df = select(babynames_df, year_of_birth, sex, ethnicity, child_s_first_name, rank) %>% 
  filter(child_s_first_name %in% c("ETHAN", "NOAH","LIAM","JOSEPH","JAYDEN","DAVID", "RYAN","MICHAEL")) %>% 
pivot_wider(names_from = year_of_birth, values_from = rank) %>% 
  relocate(child_s_first_name, ethnicity, '2016', '2015', '2014', '2013', '2012', '2011') %>% 
  select(child_s_first_name, ethnicity, '2016', '2015', '2014', '2013', '2012', '2011') %>% 
  rename("Race/Ethnicity" = ethnicity) %>% 
  rename("Rank in 2016" = '2016') %>% 
  rename("Rank in 2015" = '2015') %>% 
  rename("Rank in 2014" = '2014') %>% 
  rename("Rank in 2013" = '2013') %>% 
  rename("Rank in 2012" = '2012') %>% 
  rename("Rank in 2011" = '2011') %>% 
  arrange(child_s_first_name) %>% 
  rename("Child's Name" = child_s_first_name) %>% 
  knitr::kable()

popular_male_names_df
```

I will now create a scatter plot showing the number of children with a name vs. the rank in popularity of that name, for white non-hispanic male children born in 2016.

```{r}
scatter_plot_df = select(babynames_df, year_of_birth, sex, ethnicity, child_s_first_name, rank, count) %>% 
                         filter(year_of_birth == '2016', sex == "MALE", ethnicity == "WHITE NON HISPANIC") 

ggplot(scatter_plot_df, aes(x = rank, y = count)) + 
  geom_point() +
  xlab("Rank in Popularity of that Name") + ylab("Number of Children with Name")
```

